---
title: "Calibration of varbvs and BVSR methods"
author: "Peter Carbonetto"
date: September 5, 2017
output:
  html_document:
    theme: readable
    toc: false
    toc_float: false
    include:
      before_body: include/header.html
      after_body: include/footer.html
---

Check calibration of the varbvs and GEMMA posterior inclusion
probabilities (PIPs) in a small, simulated data set in which all the
genetic markers are uncorrelated (*i.e.*, unlinked).

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(
  eval      = TRUE,
  comment   = "#",
  results   = "hold",
  collapse  = TRUE,
  fig.align = "center")
```

Load these packages.

```{r load-pkgs, message=FALSE}
library(ggplot2)
library(cowplot)
```

Load the results of running the `demo.sim.R` script.

```{r load-results}
load("../output/demo.sim.out.RData")
```

## Compare varbvs and BVSR results

Compute the false positive rate (FPR) at each posterior inclusion
probability (PIP) threshold.

```{r compute-fpr}
compute.fpr <- function (beta, pip) {
  x <- unique(round(fit.varbvs$pip,digits = 3))
  x <- c(0,x,1)
  x <- sort(unique(x))
  n <- length(x)
  y <- rep(0,n)
  y[1] <- 1
  for (i in 2:(n-1))
    y[i] <- sum(pip >= x[i] & beta == 0) / sum(pip >= x[i])
  return(data.frame(cutoff = x,fpr = y))
}
fpr.varbvs <- compute.fpr(beta,fit.varbvs$pip)
fpr.bvsr   <- compute.fpr(beta,fit.bvsr$pip)
```

Plot FPR vs. PIP for the varbvs and BVSR methods.

```{r plot-fpr-vs-threshold, fig.width=4, fig.height=3.5}
print(ggplot() +
      geom_line(data = fpr.varbvs,mapping = aes(x = cutoff,y = fpr),
	            color = "darkorange",size = 1) +
      geom_line(data = fpr.bvsr,mapping = aes(x = cutoff,y = fpr),
	            color = "darkblue",linetype = "dashed",size = 1) +
  labs(x = "PIP threshold",y = "FPR") +
  theme_cowplot(font_size = 14))
```

## Session information

This is the version of R and the packages that were used to generate
these results in script `demo.sim.R`.

```{r session-info}
session.info
```
